An open reproducible framework for the study of the iterated prisoner's dilemma

### 2015-10-21 tex/paper.tex

As stated in~\cite{Bendor1991}: ``\textit{few works in social science have had
the general impact of [Axelrod's study of the evolution of cooperation]}''.  In
1980, Axelrod wrote two papers:~\cite{Axelrod1980a,Axelrod1980b} which
described a computer tournament that has been at the origin of a majority of
game theoretic work~\cite{Banks1990, Bendor1991, Boyd1987, Chellapilla1999,
DavidB1993, Doebeli2005, Ellison1994, Gotts2003, Hilbe2013, Isaac2008,
Kraines1989, Lee2015, Lorberbaum1994, Milgrom1982, Molander1985, Murnighan2015,
Press2012, Stephens2002, Stewart2012}. As described in~\cite{Bendor1991} this
work has not only had mathematical impact but has also led to insights in
biology (for example in~\cite{Stephens2002}, a real tournament where Blu Jays
are the participants is described) and in particular to the study of evolution.

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Numerical
values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature. Axelrod's
tournaments (and further implementations of these) are sometimes referred to as
Iterated Prisoner's Dilemma (IPD) tournaments, an overview of published tournaments is
given in Table~\ref{tab:tournaments}.

@@

In \cite{Milgrom1982} describes how incomplete information can be used to
enhance cooperation, in a similar approach to the proof of the Folk theorem for
repeated games \cite{Maschler2013}. This aspect of incomplete information is
also considered in \cite{Molander1985, Bendor1991, Lee2015} where ``noisy''
tournaments randomly flip the choice made by a given strategy. In
\cite{Murnighan2015} incomplete information is considered in the sense of a
probabilistic termination of each round of the tournament.

As mentioned before, IPD tournaments have been studied in an evolutionary
context: \cite{Ellison1994, Lee2015, Press2012, Stewart2012} consider this in a
traditional evolutionary game theory context. These works investigates
particular evolutionary contexts within which cooperation can emerge as stable.
Often these works consider direct opposition to another strategy and disregard
the population dynamics, for example in \cite{Lee2015} a simple machine learning
algorithm outperforms a strategy found in \cite{Press2012}.

Further to these evolutionary ideas, \cite{Chellapilla1999, DavidB1993} are
examples of using machine learning techniques to evolve particular strategies.
In \cite{Axelrod} Axelrod describes how similar techniques are used to
genetically evolve a high performing strategy from a given set of strategies.
Note that in his original work Axelrod only used a base strategy set of 12
strategies for this evolutionary study. This is noteworthy as
\cite{Axelrod-Pythonprojectteam2015} now boasts over 90 strategies that are
readily available for a similar analysis.

@@

Furthermore \cite{Isaac2008} Python is described as an appropriate language for the
reproduction of Iterated Prisoner's dilemma tournaments due to it's object
oriented nature and readability.

@@

As stated in Section~\ref{sec:introduction} one of the main goals of the library
is to allow for the easy contribution of strategies. To do this requires the
writing of a simple python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as it's corresponding test.

>>>>

As stated in~\cite{Bendor1991}: ``\textit{few works in social science have had
the general impact of [Axelrod's study of the evolution of cooperation]}''.  In
1980, Axelrod wrote two papers:~\cite{Axelrod1980a,Axelrod1980b} which
described a computer tournament that has been at the origin of a majority of
game theoretic work~\cite{Banks1990, Bendor1991, Boyd1987, Chellapilla1999,
DavidB1993, Doebeli2005, Ellison1994, Gotts2003, Hilbe2013, Isaac2008,
Kraines1989, Lee2015, Lorberbaum1994, Milgrom1982, Molander1985, Murnighan2015,
Press2012, Stephens2002, Stewart2012}. As described in~\cite{Bendor1991} this
work has not only had mathematical impact but has also led to insights in
biology (for example in~\cite{Stephens2002}, a real tournament where Blue Jays
are the participants is described) and in particular to the study of evolution.

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Numerical
values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature. Axelrod's
tournaments (and further implementations of these) are sometimes referred to as
Iterated Prisoner's Dilemma (IPD) tournaments. An overview of published tournaments is
given in Table~\ref{tab:tournaments}.

@@

\cite{Milgrom1982} describes how incomplete information can be used to
enhance cooperation, in a similar approach to the proof of the Folk theorem for
repeated games \cite{Maschler2013}. This aspect of incomplete information is
also considered in \cite{Molander1985, Bendor1991, Lee2015} where ``noisy''
tournaments randomly flip the choice made by a given strategy. In
\cite{Murnighan2015}, incomplete information is considered in the sense of a
probabilistic termination of each round of the tournament.

As mentioned before, IPD tournaments have been studied in an evolutionary
context: \cite{Ellison1994, Lee2015, Press2012, Stewart2012} consider this in a
traditional evolutionary game theory context. These works investigate
particular evolutionary contexts within which cooperation can emerge as stable.
Often these works consider direct opposition to another strategy and disregard
the population dynamics, for example in \cite{Lee2015} a simple machine learning
algorithm outperforms a strategy found in \cite{Press2012}.

Further to these evolutionary ideas, \cite{Chellapilla1999, DavidB1993} are
examples of using machine learning techniques to evolve particular strategies.
In \cite{Axelrod}, Axelrod describes how similar techniques are used to
genetically evolve a high performing strategy from a given set of strategies.
Note that in his original work, Axelrod only used a base strategy set of 12
strategies for this evolutionary study. This is noteworthy as
\cite{Axelrod-Pythonprojectteam2015} now boasts over 90 strategies that are
readily available for a similar analysis.

@@

Furthermore, \cite{Isaac2008} Python is described as an appropriate language for the
reproduction of Iterated Prisoner's dilemma tournaments due to its object
oriented nature and readability.

@@

As stated in Section~\ref{sec:introduction} one of the main goals of the library
is to allow for the easy contribution of strategies. To do this requires the
writing of a simple python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as its corresponding test.
